A White Noise Approach to Evolutionary Ecology

### 2020-04-13 .A White Noise Approach To Evolutionary Ecology/draft.Rmd
improved proof of classical solutions

We shift our focus to showing mild solutions $\nu$ to PDE (\ref{PDE_supp}) are twice differentiable in $x$ and hence are classical solutions. Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong solution (i.e., differentiable with respect to $t\in(0,T)$) to the abstract Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$. In particular, this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,1}(\mathbb{R})$ is the Sobolov space of integrable functions with weak derivatives at least of order $2$. Since $\frac{1}{p}\geq\frac{1}{p+1}\geq\frac{1}{p}-2$ for all $p\geq1$, the Sobolov embedding theorem [@Henry1981] implies $D(\Delta)\subset L^p(\mathbb{R})$, the Lebesgue space of $p$-integrable functions on $\mathbb{R}$, for all $p\geq1$. Hence $D(\Delta)\subset W^{2,p}(\mathbb{R})$, the Sobolov space of $p$-integrable functions with weak derivatives at least of order $2$, for all $p\geq1$.

@@

Recall that $m(\nu,x)$ is shorthand for $m\big((K\nu)(x,t),x\big)$. That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$. Following our assumptions of the main text, we have that $m(h,x)$ is differentiable with respect to both $x$ and $h$ and there exists $r\in\mathbb{R}$ such that $m(h,x)\leq r$ across all $h\geq0$ and $x\in\mathbb{R}$. As in the main text, we also assume the initial condition $\nu(x,0)$ is continuous in $x$ and satisfies
\begin{equation}
N(0)=\int_{\mathbb{R}}\nu(x,0)dx<+\infty,
\end{equation}
\begin{equation}
-\infty<\bar x(0)=\int_{\mathbb{R}}xp(x,0)dx<+\infty,
\end{equation}
\begin{equation}
\sigma^2(0)=\int_{\mathbb{R}}(x-\bar x(0))^2p(x,0)dx<+\infty,
\end{equation}
where $p(x,0)=\nu(x,0)/N(0)$, and we consider the PDE
\begin{equation}\label{PDE_SM}
\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t).
\end{equation}

In SM \S\ref{regular} we have shown PDE (\ref{PDE_SM}) with an integrable initial condition has a classical solution. Replacing $m$ with it's upper bound $r\in\mathbb{R}$, equation (\ref{PDE_SM}) reduces to a simple parabolic equation that can be solved using standard techniques [@stanleyfarlow1993]. In particular, when $m(\nu,x)\equiv0$ denote the solution to (\ref{PDE_SM}) by $\nu_0(x,t)$. Then, denoting
\begin{equation}
\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}},
\end{equation}
we have
\begin{equation}
\nu_0(x,t)=\int_{\mathbb{R}}\Phi(x-y,t)\nu(y,0)dy.
\end{equation}
In the more general case, when $m(\nu,x)\equiv r\in\mathbb{R}$, equation (\ref{PDE_SM}) has the solution $\nu_r(x,t)=e^{rt}\nu_0(x,t)$. Hence, $\nu_r(x,t)\geq0$ for all $x\in\mathbb{R}$ and $\int_\mathbb{R}\nu_r(x,t)dx=e^{rt}N(0)<+\infty$ for all $t\geq0$. Furthermore, denoting 
\begin{equation}
\begin{matrix}
N_r(t)&=&\int_\mathbb{R}\nu_r(x,t)dx, \\
p_r(x,t)&=&\nu_r(x,t)/N_r(t), \\
\bar x_r(t)&=&\int_\mathbb{R}xp_r(x,t)dx, \\
\sigma^2_r(t)&=&\int_\mathbb{R}(x-\bar x_r(t))^2p_r(x,t)dx, 
\end{matrix}
\end{equation}
we have
\begin{equation}
\bar x_r(t)=\int_\mathbb{R}x\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}yp(y,0)dy=\bar x(0),
\end{equation}
\begin{equation}
\sigma_r^2(t)=\int_\mathbb{R}(x-\bar x_r(t))^2\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}\Big((y-\bar x(0))^2+\mu t\Big)p(y,0)dy=\sigma^2(0)+\mu t.
\end{equation}
Hence, $|\bar x_r(t)|,\sigma_r^2(t)<+\infty$ for all $t\geq0$. For the sake of contradiction, suppose there exists $x\in\mathbb{R}$ and $t\geq0$ such that $\nu(x,t)>\nu_r(x,t)$. Then
\begin{equation}
\nu(x,t)-\nu(x,0)=\int_0^tm(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)ds>\int_0^tr\nu_r(x,s)+\frac{\mu}{2}\Delta\nu_r(x,s)ds=\nu_r(x,t)-\nu(x,0)
\end{equation}
which implies there exists $h\geq0$ and $x\in\mathbb{R}$ such that $m(h,x)>r$. But this contradicts our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$. So we have $\nu(x,t)\leq\nu_r(x,t)$ for each $x\in\mathbb{R}$ and $t\geq0$. This implies that $N(t)=\int_\mathbb{R}\nu(x,t)dx<+\infty$,
\begin{equation}
0\leq\int_\mathbb{R}x^2\nu(x,t)dx\leq\int_\mathbb{R}x^2\nu_r(x,t)dx<+\infty
\end{equation}
and in particular
\begin{equation}
0\leq\sigma^2(t)+\bar x^2(t)=\frac{1}{N(t)}\int_\mathbb{R}x^2\nu(x,t)dx<+\infty
\end{equation}
for each $t\geq0$.

>>>>

We shift our focus to showing mild solutions $\nu$ to PDE (\ref{PDE_supp}) are twice differentiable in $x$ and hence are classical solutions. Applying lemma 3.3.2 of Henry [-@Henry1981] we have that the mild solution $\nu$ is in fact a strong solution (i.e., differentiable with respect to $t\in(0,T)$) to the abstract Cauchy problem defined by PDE (\ref{PDE_supp}) with initial condition $\nu(0)=u$ for $t\in[0,T]$. In particular, this implies that $\nu(t)\in D(\Delta)$ for each $t\in(0,T)$ where $D(\Delta)=W^{2,1}(\mathbb{R})$ is the Sobolov space of integrable functions with integrable weak derivatives at least of order $2$. Since $\frac{1}{p}\geq\frac{1}{p+1}\geq\frac{1}{p}-2$ for all $p\geq1$, the Sobolov embedding theorem [@Henry1981] implies $D(\Delta)\subset L^p(\mathbb{R})$, the Lebesgue space of $p$-integrable functions on $\mathbb{R}$, for all $p\geq1$. Hence $D(\Delta)\subset W^{2,p}(\mathbb{R})$, the Sobolov space of $p$-integrable functions with weak derivatives at least of order $2$, for all $p\geq1$.

@@

Recall that $m(\nu,x)$ is shorthand for $m\big((K\nu)(x,t),x\big)$. That is, $m:[0,+\infty)\times\mathbb{R}\to\mathbb{R}$. Following our assumptions of the main text, we have that $m(h,x)$ is differentiable with respect to both $x$ and $h$ and there exists $r\in\mathbb{R}$ such that $m(h,x)\leq r$ across all $h\geq0$ and $x\in\mathbb{R}$. As in the main text, we also assume the initial condition $u(x)$ is continuous and integrable in $x$ and satisfies
\begin{equation}
0\leq\int_{\mathbb{R}}(|x|+x^2)u(x)dx<+\infty
\end{equation}
and consider the Cauchy problem
\begin{equation}\label{PDE_SM}
\left\{\begin{matrix}
\dot\nu(x,t)=m(\nu,x)\nu(x,t)+\frac{\mu}{2}\Delta\nu(x,t) & t>0\\
\nu(x,0)=u(x) & t=0.
\end{matrix}\right.
\end{equation}

As mentioned in the main text, theorem 2.5.6 of Zheng [-@zheng2004nonlinear] states that if $m(\nu,x)$ is locally Lipschitz in both arguments then for some maximal $T>0$ problem (\ref{PDE_SM}) admits a unique local classical solution $\nu(x,t)$ for $t\in[0,T)$. Furthermore, either $T=+\infty$ or $T<+\infty$ and $\lim_{t\uparrow T}\int_\mathbb{R}|\nu(t,x)|dx=+\infty$. In this section we show that our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$ implies $T=+\infty$. Replacing $m$ with it's upper bound $r\in\mathbb{R}$, equation (\ref{PDE_SM}) reduces to a simple parabolic equation that can be solved using elementary techniques [@stanleyfarlow1993]. In particular, when $m(\nu,x)\equiv0$ denote the solution to (\ref{PDE_SM}) by $\nu_0(x,t)$. Then, denoting
\begin{equation}
\Phi(x,t)=\frac{\exp\left(-x^2/2\mu t\right)}{\sqrt{2\pi\mu t}},
\end{equation}
we have
\begin{equation}
\nu_0(x,t)=\int_{\mathbb{R}}\Phi(x-y,t)\nu(y,0)dy.
\end{equation}
In the more general case, when $m(\nu,x)\equiv r\in\mathbb{R}$, equation (\ref{PDE_SM}) has the solution $\nu_r(x,t)=e^{rt}\nu_0(x,t)$. Hence, $\nu_r(x,t)\geq0$ for all $x\in\mathbb{R}$ and $\int_\mathbb{R}\nu_r(x,t)dx=e^{rt}N(0)<+\infty$ for all $t\geq0$. Furthermore, denoting 
\begin{equation}
\begin{matrix}
N_r(t)&=&\int_\mathbb{R}\nu_r(x,t)dx, \\
p_r(x,t)&=&\nu_r(x,t)/N_r(t), \\
\bar x_r(t)&=&\int_\mathbb{R}xp_r(x,t)dx, \\
\sigma^2_r(t)&=&\int_\mathbb{R}(x-\bar x_r(t))^2p_r(x,t)dx, 
\end{matrix}
\end{equation}
we have
\begin{equation}
\bar x_r(t)=\int_\mathbb{R}x\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}yp(y,0)dy=\bar x(0),
\end{equation}
\begin{equation}
\sigma_r^2(t)=\int_\mathbb{R}(x-\bar x_r(t))^2\int_\mathbb{R}\Phi(x-y,t)p(y,0)dydx=\int_\mathbb{R}\Big((y-\bar x(0))^2+\mu t\Big)p(y,0)dy=\sigma^2(0)+\mu t.
\end{equation}
Hence, $|\bar x_r(t)|,\sigma_r^2(t)<+\infty$ for all $t\geq0$. For the sake of contradiction, suppose there exists $x\in\mathbb{R}$ and $t\geq0$ such that $\nu(x,t)>\nu_r(x,t)$. Then
\begin{equation}
\nu(x,t)-\nu(x,0)=\int_0^tm(\nu,x)\nu(x,s)+\frac{\mu}{2}\Delta\nu(x,s)ds>\int_0^tr\nu_r(x,s)+\frac{\mu}{2}\Delta\nu_r(x,s)ds=\nu_r(x,t)-\nu(x,0)
\end{equation}
which implies there exists $h\geq0$ and $x\in\mathbb{R}$ such that $m(h,x)>r$. But this contradicts our assumption $m(h,x)\leq r$ for all $h\geq0$ and $x\in\mathbb{R}$. So we have $\nu(x,t)\leq\nu_r(x,t)$ for each $x\in\mathbb{R}$ and $t\geq0$. This implies that $N(t)=\int_\mathbb{R}\nu(x,t)dx<+\infty$,
\begin{equation}
0\leq\int_\mathbb{R}x^2\nu(x,t)dx\leq\int_\mathbb{R}x^2\nu_r(x,t)dx<+\infty
\end{equation}
and in particular
\begin{equation}
0\leq\sigma^2(t)+\bar x^2(t)=\frac{1}{N(t)}\int_\mathbb{R}x^2\nu(x,t)dx<+\infty
\end{equation}
for each $t\geq0$.
